Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube

### 2022-09-13 latex/Abstract/Abstract.tex
last cleanup

\label{Abstract}
\noindent \textbf{Motivation} -- Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, has until recently known less success than we might have hoped for. Its framework is very appealing and intuitive. In particular, the mathematical beauty of Value and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound. Sadly, it had until recently proven hard to achieve practical success with these methods. Inspired however by the seminal work of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}). I will attempt to implement a variety of solvers combining $A^{*}$ and heuristics, some handcrafted, others trained on sequences of puzzles using Deep Learning (\textbf{DL}), \textbf{DRL} or \textbf{DQL}, to solve the 15-puzzle and the Rubik's cube.
\\
\\
\textbf{Organisation of this thesis} -- In section 1, I discuss what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!). In section 2, I do a recap of the methods used in the project. Section 3 is dedicated to the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}. I give then an overview of my implementation in section 4. Sections 6 and 7 present my results on respectively the \textbf{SP} and the \textbf{RC} and section 8 concludes. A couple of appendices complement the main text: the first one details how to use the code base, the second describes a limitation I bumped into while using the kociemba library.
\\
\\
\noindent \textbf{Acknowledgements} -- 
\input{Acknowledgements/Acknowledgements}

>>>>

\label{Abstract}
\noindent \textbf{Motivation} -- Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, had until recently known less success than we might have hoped for. Its framework is very appealing and intuitive. In particular, the mathematical beauty of Value and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound. Sadly, it had proven hard to achieve practical success with these methods. Inspired however by the seminal work of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}). I will attempt to implement a variety of solvers combining $A^{*}$ and heuristics, some handcrafted, others trained on sequences of puzzles using Deep Learning (\textbf{DL}), \textbf{DRL} or \textbf{DQL}, to solve the 15-puzzle and the Rubik's cube.
\\
\\
\textbf{Organisation of this thesis} -- In section 1, I discuss what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!). In section 2, I do a recap of the methods used in the project. Section 3 is dedicated to the mathematics of the sliding puzzle (\textbf{SP}) and the \textbf{RC}. I give then an overview of my implementation in section 4. Sections 6 and 7 present my results on respectively the \textbf{SP} and the \textbf{RC} and section 8 concludes. A couple of appendices complement the main text: the first one details how to use the code base, the second describes a limitation I bumped into while using the kociemba library.
\\
\\
\noindent \textbf{Acknowledgements} -- 
\input{Acknowledgements/Acknowledgements}

### 2022-09-13 latex/Appendices/Examples.tex
last cleanup

\begin{figure}[H]
\centering
\includegraphics[scale=0.41]{./Figures/exampleperfectlearner}
%\decoRule
\caption[Examples]{perfect learner example}
\label{fig:exampleperfectlearner}
\end{figure}

@@

As can be seen in the code snippet, this example will generate 100 perfectly shuffled (n=5, m=2) SPs and solve them. Once done, a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{./Figures/exampledeeplearnertraining}
%\decoRule
\caption[Examples]{deep learner training example}
\label{fig:exampledeeplearnertraining}
\end{figure}

Then the Deep Learner will get trained on this data for 999 epochs. In the above example, I have chosen to fetch, every 100 epochs, 50 random sequences of puzzles from the training data. Each sequence is composed of a random puzzle of cost between 20 and 50, fetched from the training data, along with its optimal path to the solution. The default optimiser (rms\_prop) is used, together with an exponential scheduler starting with a learning rate of 0.001 and a gamma of 0.9999. We can see that the (MSE) loss on the value function decreases rapidly, and jumps back up every time we change the training data (since it has not yet seen some of it). By the end of the training, the in-sample MSE loss has dropped to 1.5 and the Deep Learner has seen 0.14\% of the possible (n=5, m=2) puzzles.

\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
\begin{landscape}
\centering\vspace*{\fill}
\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{./Figures/exampledeeplearnerlearning}
%\decoRule
\caption[Examples]{deep learner learning data example}
\label{fig:exampledeeplearnerlearning}
\end{figure}
\vfill
\end{landscape}
\restoregeometry

@@

Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the actual (target) costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}. Let us see here how to run it on the (n=4, m=2) SP for instance. The following code snippet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network. That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousands of the max target value. The learner will stop if it reaches 25,000 epochs or if the target network updates 40 times. The network trained is a fully connected network with 3 hidden layers and the puzzles are one-hot encoded. We use the same optimiser and scheduler as in the previous subsection.

\afblue
\paragraph{}{\textbf{python code -- deep reinforcement learner}}
\begin{python}
####################################################################
from rubiks.deeplearning.deeplearning import DeepLearning
from rubiks.learners.learner import Learner
from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
from rubiks.puzzle.puzzle import Puzzle
####################################################################
if '__main__' == __name__:
    puzzle_type=Puzzle.sliding_puzzle
    n=4
    m=2
    """ Generate training data - 100 sequences of fully 
    solved perfectly shuffled puzzles. 
    """
    nb_cpus=4
    """ DRL learner """
    action_type=Learner.do_learn
    learner_type=Learner.deep_reinforcement_learner
    nb_epochs=25000
    nb_shuffles=50
    nb_sequences=10
    training_data_every_epoch=False
    cap_target_at_network_count=True
    update_target_network_frequency=500
    update_target_network_threshold=1e-3
    max_nb_target_network_update=40
    max_target_not_increasing_epochs_pct=0.5
    max_target_uptick=0.01
    learning_rate=1e-3
    scheduler=DeepReinforcementLearner.exponential_scheduler
    gamma_scheduler=0.9999
    """ ... and its network config """
    network_type=DeepLearning.fully_connected_net
    layers_description=(128, 64, 32)
    one_hot_encoding=True
    """ Kick-off the Deep Reinforcement Learner ... """
    learning_file_name=Learner.factory(**globals()).get_model_name()
    Learner.factory(**globals()).action()
    """ ... and plot its learning data """
    action_type=Learner.do_plot
    Learner.factory(**globals()).action()
####################################################################
\end{python}
\black
As can be seen on the next page, which I obtained from running the above code snippet (keep in mind that every run is going to be slightly different due to the random puzzles being generated), the training stopped after about 17,100 epochs as the 40 target network updates had been reached. By that point, the DRL learner had seen 40\% of the possible puzzles, and the very last MSE loss (after update, so out-of-sample) was around 0.4, corresponding to 2\% of the max target cost. It is interesting to notice that since I have shuffled the sequences only 50 times, and since the (n=4, m=2) SP is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the God number for this dimension is 36 (see later section \ref{sec:SPLowDimension}). It is therefore likely that the resulting network would not produce very optimal solutions for puzzles whose cost is in the region [25, 36].

\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
\begin{landscape}
\centering\vspace*{\fill}
\begin{figure}[H]
\centering
\includegraphics[scale=0.55]{./Figures/exampledeepreinforcementlearnerlearning}
%\decoRule
\caption[Examples]{deep reinforcement learner learning data example}
\label{fig:exampledeepreinforcementlearnerlearning}
\end{figure}
\vfill
\end{landscape}
\restoregeometry

@@

%-----------------------------------
%	SUBSECTION 3.1
%-----------------------------------
\Subsection{Blind search}
First let me start with some of the blind search algorithms, which I only have been able to use on small dimension SP. They quickly become too memory hungry to be practical on anything but the smallest puzzles.
%-----------------------------------
%	SUBSECTION 3.1.1
%-----------------------------------
\Subsubsection{BFS}
\label{BFSSS}
The following example shows how to use Breadth First Search to solve a (n=3, m=3) SP which we do not \textit{shuffle} too much.

\afblue
\paragraph{}{\textbf{python code -- breadth first search solver}}
\begin{python}
####################################################################
from rubiks.solvers.solver import Solver
from rubiks.puzzle.puzzle import Puzzle
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    n=3
    m=3
    nb_shuffles=10
    solver_type=Solver.bfs
    check_optimal=True
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action())
####################################################################
\end{python}
\black

As expected (since BFS is optimal), the solution found, printed by the snippet of code above, is optimal. The check\_optimal flag in the code snippet indicates that the solver should let us know if solution is optimal. Since BFS advertises itself (via the Solver base class API)  as an optimal solver, the solution is deemed optimal.

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/examplebfssolver}
%\decoRule
\caption[Examples]{breadth first search solver example}
\label{fig:examplebfssolver}
\end{figure}

@@

\end{python}
\black
Here we can see, in contrast with the previous example, that the solution obtained by DFS is not optimal. The Solver indicates it and shows an optimal solution. Clearly since we shuffled 8 times from goal configuration, the optimal path cannot have a cost higher than 8 (it could be lower though of course).

\begin{figure}[H]
\centering
\includegraphics[scale=0.2]{./Figures/exampledfssolver}
%\decoRule
\caption[Examples]{depth first search solver example}
\label{fig:exampledfssolver}
\end{figure}

@@

\item if n and m are both equal to 2, we just keep moving the empty tile clock-wise until the puzzle is solved. Notice that this is bound to work, since moving clock-wise or counter-clock-wise are the two ony possible moves, and one of them is just un-doing the other one, therefore the only possible sequence of move in a (n=2, m=2) puzzle is to either keep moving clock-wise or counter-clock-wise.

@@

As a concrete example, let us assume we started with the following (n=6, m=6) puzzle:

@@

After one call to solve the top row and the left column, we are left with solving the (n=5,m=5) sub-puzzle in blue:

@@

and we are left with solving the bottom sub-puzzle (n=4,m=5):

@@

Below is a simple code snippet to run the naive solver on a randomly generated (n=2, m=2) SP:

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.39]{./Figures/examplenaivesolver}
%\decoRule
\caption[Examples]{naive solver example}
\label{fig:examplenaivesolver}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.55]{./Figures/examplekociemba}
%\decoRule
\caption[Examples]{Kociemba solver example}
\label{fig:examplekociemba}
\end{figure}

@@

\afblue
\paragraph{}{\textbf{python code -- depth first search solver}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    tiles=[[3, 8, 6], [4, 1, 5], [0, 7, 2]]
    solver_type=Solver.astar
    heuristic_type=Heuristic.manhattan
    plus=False
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action().to_str_light())
####################################################################
\end{python}
\black
We run this example twice, one with simple Manhattan and one with Manhattan++. As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property). See a more detailed analysis in the SP results section \ref{}

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/exampleastarmanhattansolver}
%\decoRule
\caption[Examples]{a* manhattan solver example}
\label{fig:exampleastarmanhattansolver}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/exampleastarmanhattanplussolver}
%\decoRule
\caption[Examples]{a* manhattan++ solver example}
\label{fig:exampleastarmanhattanplussolver}
\end{figure}

@@

\afblue
\paragraph{}{\textbf{python code -- depth first search solver}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.utils.utils import get_model_file_name
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    n=3
    nb_shuffles=40
    solver_type=Solver.astar
    heuristic_type=Heuristic.perfect
    model_file_name = get_model_file_name(puzzle_type=puzzle_type,
                                          dimension=(n, n),
                                          model_name=Heuristic.perfect)
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action().to_str_light())
####################################################################
\end{python}
\black
Running this code snippet will output something like (the shuffle is random obviously):

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/exampleastarperfectsolver}
%\decoRule
\caption[Examples]{a* perfect solver example}
\label{fig:exampleastarperfectsolver}
\end{figure}

\Subsubsection{Deep Learning Heuristic}
\Subsubsection{Deep Reinforcement Learning Heuristic}
\Subsubsection{Deep Q Learning Heuristic}

>>>>

\begin{figure}[H]
\centering
\includegraphics[scale=0.41]{./Figures/exampleperfectlearner}
\caption[Examples]{perfect learner example}
\label{fig:exampleperfectlearner}
\end{figure}

@@

\noindent As can be seen in the code snippet, this example will generate 100 perfectly shuffled 5x2 \textbf{SP}s and solve them. Once done, a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{./Figures/exampledeeplearnertraining}
\caption[Examples]{deep learner training example}
\label{fig:exampledeeplearnertraining}
\end{figure}

\noindent Then the Deep Learner will get trained on this data for 999 epochs. In the above example, I have chosen to fetch, every 100 epochs, 50 random sequences of puzzles from the training data. Each sequence is composed of a random puzzle of cost between 20 and 50, fetched from the training data, along with its optimal path to the solution. The default optimiser (rms\_prop) is used, together with an exponential scheduler starting with a learning rate of 0.001 and a gamma of 0.9999. We can see that the (MSE) loss on the value function decreases rapidly, and jumps back up every time we change the training data (since it has not yet seen some of it). By the end of the training, the in-sample MSE loss has dropped to 1.5 and the Deep Learner has seen 0.14\% of the possible 5x2 puzzles.

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/exampledeeplearnerlearning}}
  \caption[Examples]{deep learner learning data example}
  \label{fig:exampledeeplearnerlearning}
\end{figure}

@@

Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the actual (target) costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}. Let us see here how to run it on the 4x2 \textbf{SP} for instance. The following code snippet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network. That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousandth of the max target value. The learner will stop if it reaches 25,000 epochs or if the target network updates 40 times. The network trained is a fully connected network with 3 hidden layers and the puzzles are one-hot encoded. We use the same optimiser and scheduler as in the previous subsection.

\afblue
\paragraph{}{\textbf{python code -- deep reinforcement learner}}
\begin{python}
####################################################################
from rubiks.deeplearning.deeplearning import DeepLearning
from rubiks.learners.learner import Learner
from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
from rubiks.puzzle.puzzle import Puzzle
####################################################################
if '__main__' == __name__:
    puzzle_type=Puzzle.sliding_puzzle
    n=4
    m=2
    """ Generate training data - 100 sequences of fully 
    solved perfectly shuffled puzzles. 
    """
    nb_cpus=4
    """ DRL learner """
    action_type=Learner.do_learn
    learner_type=Learner.deep_reinforcement_learner
    nb_epochs=25000
    nb_shuffles=50
    nb_sequences=10
    training_data_every_epoch=False
    cap_target_at_network_count=True
    update_target_network_frequency=500
    update_target_network_threshold=1e-3
    max_nb_target_network_update=40
    max_target_not_increasing_epochs_pct=0.5
    max_target_uptick=0.01
    learning_rate=1e-3
    scheduler=DeepReinforcementLearner.exponential_scheduler
    gamma_scheduler=0.9999
    """ ... and its network config """
    network_type=DeepLearning.fully_connected_net
    layers_description=(128, 64, 32)
    one_hot_encoding=True
    """ Kick-off the Deep Reinforcement Learner ... """
    learning_file_name=Learner.factory(**globals()).get_model_name()
    Learner.factory(**globals()).action()
    """ ... and plot its learning data """
    action_type=Learner.do_plot
    Learner.factory(**globals()).action()
####################################################################
\end{python}
\black
As can be seen on the next page, which I obtained from running the above code snippet (keep in mind that every run is going to be slightly different due to the random puzzles being generated), the training stopped after about 17,100 epochs as the 40 target network updates had been reached. By that point, the DRL learner had seen 40\% of the possible puzzles, and the very last MSE loss (after update, so out-of-sample) was around 0.4, corresponding to 2\% of the max target cost. It is interesting to notice that since I have shuffled the sequences only 50 times, and since the 4x2 \textbf{SP} is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the God number for this dimension is 36 (see later section \ref{sec:SPLowDimension}). It is therefore likely that the resulting network would not produce very optimal solutions for puzzles whose cost is in the region [25, 36].

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/exampledeepreinforcementlearnerlearning}}
  \caption[Examples]{deep reinforcement learner learning data example}
  \label{fig:exampledeepreinforcementlearnerlearning}
\end{figure}

@@

%-----------------------------------
%	SUBSECTION 3.1
%-----------------------------------
\Subsection{Blind search}
First let me start with some of the blind search algorithms, which I only have been able to use on small dimension SP. They quickly become too memory hungry to be practical on anything but the smallest puzzles.
%-----------------------------------
%	SUBSECTION 3.1.1
%-----------------------------------
\Subsubsection{BFS}
\label{BFSSS}
The following example shows how to use Breadth First Search to solve a 3x3 \textbf{SP} which we do not \textit{shuffle} too much.

\afblue
\paragraph{}{\textbf{python code -- \textbf{BFS} solver}}
\begin{python}
####################################################################
from rubiks.solvers.solver import Solver
from rubiks.puzzle.puzzle import Puzzle
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    n=3
    m=3
    nb_shuffles=5
    solver_type=Solver.bfs
    check_optimal=True
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action())
####################################################################
\end{python}
\black

As expected (since \textbf{BFS} is optimal), the solution found, printed by the snippet of code above, is optimal. The check\_optimal flag in the code snippet indicates that the solver should let us know if solution is optimal. Since BFS advertises itself (via the Solver base class API)  as an optimal solver, the solution is deemed optimal.

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.32]{./Figures/examplebfssolver}}
  \caption[Examples]{breadth first search solver example}
  \label{fig:examplebfssolver}
\end{figure}

@@

\end{python}
\black
Here we can see, in contrast with the previous example, that the solution obtained by \textbf{DFS} is not optimal. The Solver indicates it and shows an optimal solution. Clearly since we shuffled 8 times from goal configuration, the optimal path cannot have a cost higher than 8 (it could be lower though of course).

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.22]{./Figures/exampledfssolver}}
  \caption[Examples]{depth first search solver example}
  \label{fig:exampledfssolver}
\end{figure}

@@

\item if n and m are both equal to 2, we just keep moving the empty tile clock-wise until the puzzle is solved. Notice that this is bound to work, since moving clock-wise or counter-clock-wise are the two ony possible moves, and one of them is just un-doing the other one, therefore the only possible sequence of move in a 2x2 puzzle is to either keep moving clock-wise or counter-clock-wise.

@@

As a concrete example, let us assume we started with the following 6x6 puzzle:

@@

After one call to solve the top row and the left column, we are left with solving the 5x5 sub-puzzle in blue:

@@

and we are left with solving the bottom 4x5 sub-puzzle:

@@

Below is a simple code snippet to run the naive solver on a randomly generated 2x2 SP:

@@

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.42]{./Figures/examplenaivesolver}}
  \caption[Examples]{naive solver example}
  \label{fig:examplenaivesolver}
\end{figure}

@@

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.58]{./Figures/examplekociemba}}
  \caption[Examples]{Kociemba solver example}
  \label{fig:examplekociemba}
\end{figure}

@@

\afblue
\paragraph{}{\textbf{python code -- A$^{*}$ solver}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    tiles=[[3, 8, 6], [4, 1, 5], [0, 7, 2]]
    solver_type=Solver.astar
    heuristic_type=Heuristic.manhattan
    plus=False
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action().to_str_light())
####################################################################
\end{python}
\black
We run this example twice, one with simple Manhattan and one with Manhattan++. As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property). See a more detailed analysis in the \textbf{SP} results section \ref{MHComp}

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/exampleastarmanhattansolver}}
  \caption[Examples]{A$^{*}$[Manhattan] solver example}
  \label{fig:exampleastarmanhattansolver}
\end{figure}

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/exampleastarmanhattanplussolver}}
  \caption[Examples]{A$^{*}$[Manhattan++] solver example}
  \label{fig:exampleastarmanhattanplussolver}
\end{figure}

@@

\afblue
\paragraph{}{\textbf{python code -- A$^{*}$[Perfect] solver}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.utils.utils import get_model_file_name
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    n=3
    nb_shuffles=40
    solver_type=Solver.astar
    heuristic_type=Heuristic.perfect
    model_file_name = get_model_file_name(puzzle_type=puzzle_type,
                                          dimension=(n, n),
                                          model_name=Heuristic.perfect)
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action().to_str_light())
####################################################################
\end{python}
\black
Running this code snippet will output something like (the shuffle is random obviously):

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/exampleastarperfectsolver}}
  \caption[Examples]{A$^{*}$[Perfect] solver example}
  \label{fig:exampleastarperfectsolver}
\end{figure}

### 2022-09-13 latex/Appendices/Kociemba.tex
update
last cleanup

\chapter{Kociemba Bug} % Main chapter title

@@

Let me discuss in this appendix in more details the issue I mentioned in section \ref{sec:ResRubiks333} regarding the \textit{poor} interface of the kociemba (3x3x3) library.
\\
\\
I personally like to follow the great Scott Meyers' advice (see item 18 of \cite{meyers2005effective}), which is to not merely make interfaces easy to use, but also difficult to misuse. In that instance, I would argue they have made it very easy to shoot oneself in the foot with their interface which assumes the cubes passed to the solver are in \textit{standard} form (my terminology), not only without making that explicit anywhere, but worse than that: sometimes non standard initial configuration raise an exception, sometimes not and returning a solution which is not really one! So what is the problem exactly?
\\
\\
To start with, let us remember from chapter \ref{sec:Puzzles} that each \textbf{RC} configuration has 24 equivalent configurations under invariance by full rotation of the cube in space. Out of 24 equivalent configurations, and for an observer facing the cube, there is only one whose center front (F) cubie is red and center up (U) cubie is white.
\\
\\
In hkociemba (the third parties 2x2x2 implementation which I used under the hood of my KociembaSolver), the problem of these equivalent puzzles is irrelevant (since full cube rotation can always be achieved by 2 rotations of 2 opposite faces in opposite directions, e.g. FB' is equivalent to full cube rotation around F (CF in my jargon)). That being said, the hkociemba is smarter than just using the equivalence between full cube rotation and two faces' rotations. It does automatic color scheme recognition when passed a \textit{cube string}, and gets a smart solution given that coloring scheme. For instance, if we pass it a solved configuration, but not in standard form, it knows there is no need to do anything. In the general case of a scrambled cube, it will also endeavour to find solutions to the closest among the 24 equivalent solved configurations. Let's look at the following concerete example:

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug222}
%\decoRule
\caption[Kociemba Bug]{2x2x2 \textbf{RC} -- hkociemba finds path to closest of the 24 equivalent goals}
\label{fig:KociembaBug222}
\end{figure}
As we can see, hkociemba happily returned a 1-move solution R to green F, white U, instead of the costlier solution it would have taken to get to \textit{standard} form red F white U (clearly at a cost of 3 with solution RUD').
\\
\\
In the 3x3x3 case, things are not quite that simple. There is no way to reproduce full cube rotation from faces rotations alone, since the center cubies will never move via the latter, but do via the former. If we run the equivalent of the above problem with the (3x3x3) kociemba library, we silently get to a false solution as is apparent from printing the resulting cube.
\afblue
\paragraph{}{\textbf{Kociemba bug -- 3x3x3 \textbf{RC} -- example of \textbf{bad} interface design}}
\begin{python}
#####################################################################
from kociemba import solve
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=3
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
print(cube)
cube_string = to_kociemba(cube)
solution = solve(cube_string)
print('Kociemba Solution:', solution)
print(cube.apply_moves(from_kociemba(solution)))
###################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug333}
%\decoRule
\caption[Kociemba Bug]{3x3x3 \textbf{RC} -- kociemba does not check center cubies and returns wrong solution}
\label{fig:KociembaBug333}
\end{figure}
In order to counteract that, and make the comparisons between Kociemba 3x3x3 and other solvers and methods fairer, I have had to implement a layer on top of kociemba. My Kociemba solver basically will check if a cube is in \textit{standard} form, and if not, it will find the sequence of full cube rotations that bring us to an equivalent configuration in \textit{standard} form, pass that to equivalent \textit{standard} cube to kociemba, and stich the cube rotation to the kociemba solution (treating cube rotation as 0 cost everywhere in the code). That way, we seamlessly get to an answer that makes sense from kociemba, without the user having to worry about whether or not the cube they are passing as input is in \textit{standard} form or not! Running the same example as above via my KociembaSolver, we get the expected result, composed of 3 cube rotations to bring it to \textit{standard} form, followed by one face rotation to solve it:

>>>>

\Appendix{Kociemba Bug} % Main chapter title

@@

Let me discuss in this appendix in more details the issue I mentioned in section \ref{sec:ResRubiks333} regarding the \textit{poor} interface of the kociemba (3x3x3) library.
\\
\\
I personally like to follow the great Scott Meyers' advice (see item 18 of \cite{meyers2005effective}), which is to not merely make interfaces easy to use, but also difficult to misuse. In that instance, I would argue they have made it very easy to shoot oneself in the foot with their interface which assumes the cubes passed to the solver are in \textit{standard} form (my terminology), not only without making that explicit anywhere, but worse than that: sometimes nonstandard initial configuration raise an exception, sometimes not and returning a solution which is not really one! So what is the problem exactly?
\\
\\
To start with, let us remember from chapter \ref{sec:Puzzles} that each \textbf{RC} configuration has 24 equivalent configurations under invariance by full rotation of the cube in space. Out of 24 equivalent configurations, and for an observer facing the cube, there is only one whose centre front (F) cubie is red and centre up (U) cubie is white.
\\
\\
In hkociemba (the third parties 2x2x2 implementation which I used under the hood of my KociembaSolver), the problem of these equivalent puzzles is irrelevant (since full cube rotation can always be achieved by 2 rotations of 2 opposite faces in opposite directions, e.g. FB' is equivalent to full cube rotation around F (CF in my jargon)). That being said, the hkociemba is smarter than just using the equivalence between full cube rotation and two faces' rotations. It does automatic color scheme recognition when passed a \textit{cube string}, and gets a smart solution given that coloring scheme. For instance, if we pass it a solved configuration, but not in standard form, it knows there is no need to do anything. In the general case of a scrambled cube, it will also endeavour to find solutions to the closest among the 24 equivalent solved configurations. Let's look at the following concrete example:

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug222}
%\decoRule
\caption[Kociemba Bug]{2x2x2 \textbf{RC} -- hkociemba finds path to closest of the 24 equivalent goals}
\label{fig:KociembaBug222}
\end{figure}
As we can see, hkociemba happily returned a 1-move solution R to green F, white U, instead of the costlier solution it would have taken to get to \textit{standard} form red F white U (clearly at a cost of 3 with solution RUD').
\\
\\
In the 3x3x3 case, things are not quite that simple. There is no way to reproduce full cube rotation from faces rotations alone, since the centre cubies will never move via the latter but do via the former. If we run the equivalent of the above problem with the (3x3x3) kociemba library, we silently get to a false solution as is apparent from printing the resulting cube.
\afblue
\paragraph{}{\textbf{Kociemba bug -- 3x3x3 \textbf{RC} -- example of \textbf{bad} interface design}}
\begin{python}
#####################################################################
from kociemba import solve
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=3
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
print(cube)
cube_string = to_kociemba(cube)
solution = solve(cube_string)
print('Kociemba Solution:', solution)
print(cube.apply_moves(from_kociemba(solution)))
###################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug333}
%\decoRule
\caption[Kociemba Bug]{3x3x3 \textbf{RC} -- kociemba does not check centre cubies and returns wrong solution}
\label{fig:KociembaBug333}
\end{figure}
In order to counteract that, and make the comparisons between Kociemba 3x3x3 and other solvers and methods fairer, I have had to implement a layer on top of kociemba. My Kociemba solver basically will check if a cube is in \textit{standard} form, and if not, it will find the sequence of full cube rotations that bring us to an equivalent configuration in \textit{standard} form, pass that to equivalent \textit{standard} cube to kociemba, and stich the cube rotation to the kociemba solution (treating cube rotation as 0 cost everywhere in the code). That way, we seamlessly get to an answer that makes sense from kociemba, without the user having to worry about whether or not the cube they are passing as input is in \textit{standard} form or not! Running the same example as above via my KociembaSolver, we get the expected result, composed of 3 cube rotations to bring it to \textit{standard} form, followed by one face rotation to solve it:
